# Assignment 2 - Course 5 in Data Science Specialization

Load data
```{r loaddata}
wd = "C:/Users/areddington/OneDrive - Deloitte (O365D)/R_dir_DS_specialization/Course5/Ass2"
setwd(wd)

data <- read.csv("activity.csv", header = TRUE)
```

## Question 1: What is the mean total number of steps taken per day?
Mean total number of steps calculation (ignore missing values)
```{r q1}
mean_steps <- with(data, tapply(steps, as.factor(data$date), sum, na.rm = TRUE))
```

Histogram
```{r histo}
library(ggplot2)
qplot(mean_steps, main = "Total number of steps per day", xlab="Number of steps", ylab='Frequency', binwidth=1500)

```

A summary (including mean and median) of the total number of steps taken per day:
```{r meanetc}
summary(mean_steps)
```

## Question 2: What is the average daily activity pattern?

Get rid of missing values
```{r actpattern}
data_nona <- data[!is.na(data$steps),]
avg_steps_per_timestep <- with(data_nona, tapply(steps, data_nona$interval, mean))
time_steps <- levels(as.factor(data_nona$interval))

```

time series plot
```{r plotsteps}
plot(time_steps, avg_steps_per_timestep, type='l', col=1, main = "Average daily activity pattern", xlab = "Time intervals", ylab = "Number of steps taken on average")
```

The maximum steps are taken at this interval:
```{r maxsteps}
maxsteps = which.max(avg_steps_per_timestep)
maxsteps[1]
```

## Imputing missing values

Calculate and report the total number of missing values in the dataset
```{r na}
length(which(is.na(data$steps)))
```

Strategy for filling in all of the missing values in the dataset & create a new dataset that is equal to the original dataset but with the missing data filled in
```{r newset}
meansteps <- with(data_nona, tapply(steps, data_nona$interval, mean))
data2 <- data[is.na(data$steps),]
data2$steps <- meansteps

data_imputed <- rbind(data_nona, data2)
# data_imputed <- data_imputed[order(data_imputed$date), ]

```
Make histogram of number of steps for day with new data set
```{r hist}
steps2 <- with(data_imputed, tapply(steps, as.factor(data_imputed[order(data_imputed$date), ]$date), sum))
library(ggplot2)
qplot(steps2, main = "Total number of steps per day", xlab="Number of steps", ylab='Frequency', binwidth=1500)
```
Report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment?

Summary of original data (including mean and median):
```{r report}
summary(mean_steps)
```
Summary of imputed data (including mean and median):
```{r report2}
summary(steps2)
```
The effect of imputing the data on the mean is that both the mean and median of the data are increased.


## Are there differences in activity patterns between weekdays and weekends?

Making a new factor variable in the imputed data with two levels: Weekday and weekend
```{r detweek}
data_imputed$dateType <-  ifelse(as.POSIXlt(data_imputed$date)$wday %in% c(0,6), 'weekend', 'weekday')
```
Panel plot containing a time series plot of steps taken on weekdays or during the weekend
```{r plotpanel}
stepsimpute <- aggregate(steps ~ interval+dateType, data_imputed, mean)

gg <- ggplot(stepsimpute, aes(interval, steps)) +
    geom_line(stat = "identity", aes(colour = dateType)) +
    theme_gray() +
    facet_grid(dateType ~ ., scales="fixed", space="fixed") +
    labs(x="Interval", y=expression("No of Steps")) +
    ggtitle("No of steps Per Interval by day type")
print(gg)
```
